Linear Stochastic Approximation- Constant Step-Size and Iterate Averaging

### 2017-05-08 instance.tex
updated version

\section{Instance Dependent Bounds}
In this section, we derive instance dependent bounds for $L_t$ in \eqref{eq:exploss} which is valid for a fixed $P$ and in the next section, explore the question of deriving uniform bounds for all $P\in \P$ (where $\P$ is a class of distributions over $\C^{\ncn}$). We first present the results and then follow it up with a discussion.
\begin{lemma}\label{lm:pd}
Let $P$ be a positive definite distribution over $\C^{\ncn}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
\end{lemma}
\begin{theorem}\label{th:pdrate}
Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C^{\ncn}$. Then
\begin{align}
\EE{\norm{\thh_t-\theta^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}}
\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\theta^*}}{t+1} \right)\,.
%\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
\end{align}

@@

Let $P$ be as in \Cref{lsadist}. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, and suppose if $P_U$ is positive definite, then

@@

@@

\EE{\norm{\thh_t-\theta^*}^2}

@@

\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U}^2\norm{U^{-1}}^2}{\alpha \rhos{P_U}} \left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.

@@

\subsection{Results Discussion}
\textbf{Error Recursion and Contraction Maps:}
The output $\thh_t$ of the algorithm \eqref{eq:lsa} is the average of the internal states at times $s=0,\ldots,t-1$. Hence, before we proceed to interpret the results in \Cref{lm:pd,thm:pdrate,lm:simtran,thm:simtran,cor:as}, it is worthwhile to spend some effort in trying to understand the dynamics of the internal states by looking at the behavior of $e_t\eqdef \theta_t-\ts$.

@@

Notice that \eqref{eq:err} places our definitions of $\rhod{P}$ and $\rhos{P}$ in \Cref{def:contract}. Here the subscripts $s$ and $d$ stand for \emph{stochastic} and \emph{deterministic}, since they are associated with the spectral radius of stochastic and deterministic matrices respectively.\\
\textbf{Positive Definiteness:} The first result in \Cref{lm:pd} shows if $P$ is PD, then there exists a stepsize for which $\rhos{P}, \rhod{P}>0$ which in turn implies that the spectral radii in \eqref{eq:spectralrand} are less than $1$.
\textbf{Similarity Transformation:} 
\paragraph{Bias and Variance:} The means-squared error at $t$ is bounded by a sum of two terms. The first term is the bias term given by $\B=(1+4(\alpha\rho_{\alpha})^{-1}) (\alpha\rho_{\alpha})^{-1} \Big(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}\Big)$.  The \emph{Bias} term that captures the rate at which the initial condition $\norm{\theta_t-\ts}^2$ is forgotten. The second term is the \emph{Variance} given by $\V=(1+4(\alpha\rho_{\alpha})^{-1}) (\alpha\rho_{\alpha})^{-1} \Big(\frac{\alpha^2\sigma^2}{t+1} \Big)$. The variance term that captures the rate at which noise is rejected. The $\B$ and $\V$ terms capture two different sources of errors. To see this, note that in the absence of noise i.e., when $\zeta_t=0,~\forall t\geq 0$, we have $\E[\norm{H\eb_t}^2]=\B$ and in the presence of noise, i.e., $\zeta\neq 0,\forall t\geq 0$ and the perfect initial condition, i.e., $\theta_t=\ts$, we have $\E[\norm{H\eb_t}^2]=\V$.
\paragraph{Behaviour for extreme value of $\alpha$:} For smaller values of stepsize, i.e., $\alpha\approx 0$, the bias term blows up, due to the presence of $\alpha^{-1}$ term. This is due to the fact that the stepsizes determine the learning rate and for smaller stepsizes the learning rate is slower. However, in this case the noise term does not blow up, a fact that can appreciated by looking at \eqref{lsergen} where $\alpha$ is seen to multiply the noise term $\zeta_t$. In quantitative terms, we can see that the $\alpha^{-2}$ and $\alpha^2$ terms can each other. For larger values of $\alpha$ i.e., $\alpha\ra \alpha_{\max}$, the bounds blow up again, due to the fact that $\rho_{\alpha}\ra 0$ in this case. This is due to the fact that the effect of both noise and initial conditions decay with the contraction factor of $F_{t,i}$, which gets closer to unity as $\alpha\ra\alpha_{\max}$.

>>>>

\section{Instance Dependent Bounds}\label{sec:mainresults}
In this section, we derive instance dependent bounds  which is valid for a fixed $P$ and in the \Cref{sec:uniform}, address the question of deriving uniform bounds $\forall\,P\in \P$, where $\P$ is a class of distributions over $\C^{\dcd}$. We first present the results and then follow it up with a discussion.
\begin{lemma}\label{lm:pd}
Let $P$ be a positive definite distribution over $\C^{\dcd}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
\end{lemma}
\begin{comment}
\begin{proof}
\begin{align*}
\rhos{P}&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^*\EE{A_t^* A_t} x\\
&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^* A^*_P A_P -\alpha x^* \EE{M_t^* M_t} x\\
&\geq \lambda_{\min}(A^*_P+A_P)-\alpha \norm{A_P}^2-\sigma^2_P
\end{align*}
The proof is complete by choosing $\alpha_P<\frac{\lambda_{\min}(A^*_P+A_P)}{\norm{A_P}^2+\sigma^2_P}$
\end{proof}
\end{comment}
\begin{lemma}\label{lm:simtran}[Similarity Transformation]
Let $P^{AS}$ be asymptotically stable, then there exists a $U\in \gln$ such that $P^{AS}_U$ is positive definite.
\end{lemma}

\begin{theorem}\label{th:pdrate}
Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be a positive definite distribution over $\C^{\dcd}$. Then
\begin{align*}
\EE{\norm{\thh_t-\ts}^2}
\leq
\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}}
\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
%\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\ts})}{t+1} \right)\,.
\end{align*}

@@

Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be AS. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\ts$, and suppose if $P_U$ is positive definite, then
\begin{align*}

@@

\end{align*}

@@

\begin{align*}
\EE{\norm{\thh_t-\ts}^2}

@@

\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}}
\end{align*}

@@

\subsection{Result Discussion}
\begin{itemize}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
\item \textbf{Role of $\rhos{P}$ and $\rhod{P}$:} We can expand the MSE as \begin{align}\label{eq:exp}\EE{\norm{\eh_t}}=\frac{1}{(t+1)^2}\ip{\sum_{s=0}^t e_s,\sum_{s=0}^t e_s}.\end{align} The inner product in \eqref{eq:exp} is a summation of \emph{diagonal} terms $\EE{\ip{e_i,e_i}}$ and \emph{corss} terms of$\EE{\ip{e_i,e_j}},i\neq j$. The growth of the diagonal terms and the cross terms depends on the spectral norm of the matrices $I-\alpha A_t$ and $I-\alpha A_P$ respectively. These are given by
\begin{comment}
The output $\thh_t$ of the algorithm \eqref{eq:lsa} is the average of the internal states at times $s=0,\ldots,t-1$. The error dynamics of the internal states by looking at the behavior of $e_t\eqdef \theta_t-\ts$.

@@

Note that \eqref{eq:spectralrand} explicitly connects $\rhod{P}$ and $\rhos{P}$ in \Cref{def:contract} to the spectral norms of the aforementioned matrices. For the MSE to be bounded, we need the spectral norms to be less than unity, which in turn translates to the requirement that $\rhos{P}>0$ and $\rhod{P}>0$.
\item \textbf{Role of $U$:} When $A_P$ is positive definite, it is straighforward to produce an $\alpha>0$ such that $\rhos{P}>0$ and $\rhod{P}>0$. However, $A_P$ is Hurwitz does not guarantee positive definiteness (see \Cref{ex:pdas}) and a result an $\alpha>0$ such that $\rhos{P}>0$ and $\rhod{P}>0$. This poses a difficulty in applying results of \Cref{th:pdrate} in the case Hurwitz matrices. This can be remedied by a similiarity transformation using an appropriate $U$ that transforms $A_P$ to appropriate Jordon form $A_{P_U}$. \Cref{th:simtran} presents the rates for the transformed iterates   $\gamma_t=U^{-1 }\theta_t$. Note that \Cref{cor:as} is obtained from \Cref{th:simtran} by the fact that $\theta_t=U\gamma_t$.
\begin{example}\label{ex:pdas}
Matrices $A_1=\begin{bmatrix} 2 &1 \\-1 &0\end{bmatrix}$$A_2=\begin{bmatrix}1 & -1\\ 1 &0 \end{bmatrix}$ and $A_3=\begin{bmatrix} 1 & 10-\epsilon \\ 0 &1 \end{bmatrix}$ are not positive definite. However choosing $U_1=\begin{bmatrix} \frac{1}{\sqrt{2}}& -\frac{1}{\sqrt{2}}\\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}} \end{bmatrix}$, $U_2=\begin{bmatrix} \frac{\sqrt{2}}{1-i\sqrt{3}} & \frac{\sqrt{2}}{1+i\sqrt{3}} \\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}}\end{bmatrix}$ and $U_3=\begin{bmatrix} 1 & 0\\ 0 &\frac{1}{10} \end{bmatrix}$, we can see that matrices $J_i=U_i^{-1}A_i U_i,i=1,2,3$ are positive definite.
\end{example}
\item \textbf{Bias and Variance:} The means-squared error at $t$ is bounded by a sum of two terms. The first term is the bias term given by $\B=\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}}\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}$.  The \emph{bias} term that captures the rate at which the initial condition $\norm{\theta_0-\ts}^2$ is forgotten. The second term is the \emph{variance} given by $\V=\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}} \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} $. The variance term that captures the rate at which noise is rejected.
\item \textbf{$\alpha$, $\rhos{P}$ and $\rhod{P}$} only affect the constants in the rates and not the exponents $\frac{1}{t}$ for variance and $\frac{1}{t^2}$ for bias terms.
\item \textbf{Behaviour for extreme value of $\alpha$:} For smaller values of stepsize, i.e., $\alpha\approx 0$, the bias term blows up, due to the presence of $\alpha^{-1}$ term. This is due to the fact that the stepsizes determine the learning rate and for smaller stepsizes the learning rate is slower. However, in this case the noise term does not blow up, a fact that can appreciated by looking at \Cref{th:pdrate,th:simtran,cor:as} where $\alpha$ is seen to multiply the variances $\sigma^2_P$ and $\sigma^2_b$. In quantitative terms, we can see that the $\alpha^{-2}$ and $\alpha^2$ terms can each other. For larger values of $\alpha$ i.e., $\alpha\ra \alpha_{P}$ (or $\alpha_{P_U}$ as the case may be), the bounds blow up again, due to the fact that $\rhos{P}\ra 0$ in this case.
\item \textbf{Geometric Series:} In simple terms the error dynamics is akin to sum of a geometric series with common factor $\alpha\rho$, a fact that is illustrated by the simple $1$-dimensional example in \Cref{ex:geo}.
\begin{example}\label{ex:geo}
Consider the one dimensional case where $A_t=\rho$ for some $0<\rho<1$, $b_t$ a bounded $i.i.d$ zero mean random variable with variance $\sigma^2$, and $\ts=0$. We know that $e_t=(1-\alpha\rho)^t e_0+\sum_{s=1}^t (1-\alpha \rho)^{t-s}\alpha b_s$ and the MSE is given by
\begin{align*}
\EE{\norm{\theta_t}^2}&=\frac{1}{(t+1)^2}\left( (\alpha\rho)^{-1}(1-\alpha\rho)^{t+1}) \theta_0+ \sum_{s=1}^t (\alpha\rho)^{-1}(1-\alpha \rho)^{t-s+1} \alpha b_s\right)^2\\
&\leq (\alpha\rho)^{-2}\left(\frac{\norm{\theta_0}^2}{t^2}+\frac{\alpha^2\sigma^2}{t}\right)
\end{align*}
\end{example}
\item \textbf{Lower Bounds:}
Consider the LSA with $b_t=0$, and $P^0$ has all the mass concentrated on $A_{P^0}=\begin{bmatrix} \lambda_{\min} &0\\ 0& \lambda_{\max}\end{bmatrix}$, for some $\lambda_{\max}>\lambda_{\min}>0$. Note that in this example $\ts=0$.
By choosing $\alpha<\frac2{\lambda_{\max}}$, in this case it is straightforward to write the expression for $\eh_t$ explicitly as below:
\begin{align*}
\eh_t&=\frac{1}{t+1}\sum_{s=0}^t e_t = \frac{1}{t+1}\sum_{s=0}^t (I-\alpha A_P)^{t-s} e_0\\
&=\frac{1}{t+1}(\alpha A)^{-1}\left(I-(I-\alpha A)^{t+1}\right)e_0
\end{align*}
\begin{comment}
\begin{align*}
\norm{\eh_t}^2&=\frac{1}{(t+1)^2}\norm{(\alpha A_P)^{-1}\left(I-(I-\alpha A_P)^{t+1}\right)e_0}^2
&\leq\frac{1}{(t+1)^2}\alpha^2\norm{A_P^{-1}}^2\norm{e_0}^2
\end{align*}
\end{comment}

\begin{align*}
\norm{\eh_t}^2\geq \eh^2_t(1)=\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)
\end{align*}
Note that in this example, $\rhos{P^0}=\rhod{P^0}=\lambda_{\min} -\alpha \lambda_{\min}^2=\lambda_{\min}(1-\alpha \lambda_{\min})$. Sandwiching the error between the lower and upper bounds we have
\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)\leq \norm{\eh_t}^2\leq
\frac{1}{(t+1)^2} \left(\alpha^{-1}\rhos{P^0}+\alpha^{-2}\rhos{P^0}^2\right)\norm{\theta_0}^2
\begin{example}
Consider distributions $P^1$, $P^2$ and $P^3$ which are concentrated on matrices $A_{P^1}=\begin{bmatrix} 10 &100\\ 0 &10\end{bmatrix}$, $A_{P^2}=\begin{bmatrix} 10 &10\\ 0 &10\end{bmatrix}$, $A_{P^3}=\begin{bmatrix} 10 &0 \\ 0 &10\end{bmatrix}$. Note that $P^1$, $P^2$ and $P^3$ are respectively AS, PD, SPD respectively. Consider the LSA for the three case of $P^1$, $P^2$ and $P^3$ with $b_t=0$ and $\ts=0$
We now list the following observations:
\begin{itemize}%[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
\item In all the three instance, $0<\alpha<\frac2{10}$ ensures that the error $\norm{\eh_t}\ra 0$ as $t\ra\infty$. Thus the range of $\alpha$ is uniform across $P^i,i=1,2,3$.
\item $\rhos{P^1}=\rhod{P^1}<0$, and hence one cannot apply results of \Cref{th:pdrate}. However, with  $U=\begin{bmatrix}1 &0 \\ 0 &0.2-\epsilon\end{bmatrix}$ for some $\epsilon>0$ we have $P^1_U$ to be the distribution concentrated on the matrix $A_{P^1_U}=\begin{bmatrix} 10 &20 \\ 0 &10\end{bmatrix}$ and one can apply the results of \Cref{thm:simtran,cor:as}. Notice that the condition number $\norm{U}{U^-1}\approx 5$.
\end{itemize}
We plot $L_t$ for the three cases for $t=1,\ldots,100$ and we can see that the condition number $\norm{U}{U^-1}\approx 5$, shows up in the AS case ($P^1$).
\end{example}
\end{itemize}
\paragraph{Related Work}
LSA with CS-PR was considered in \cite{polyak-judisky} under the additive noise assumption, wherein $A_t=A$. Recent work by \cite{bach} considers a stochastic gradient descent (SGD) algorithm for the problem of linear prediction given in \Cref{ex:linreg}.
\begin{example}\label{ex:linreg}
Let $(x_t,y_t)\in \R^d\times \R,\,t\geq 1$ be $i.i.d$ such that $\EE{\norm{x_t}^2}$ and $\EE{y_t^2}$ are finite. Further, we assume bounded data i.e., $\norm{(x_t,y_t)}^2\leq B$ for some positive real number $B>0$. Here, $x_t,\,t\geq 1$ are the features and $y_t$ are real labels and the linear regression problem constitutes minimizing the quadratic loss function $f(\theta)=\EE{\left(\ip{x_t,\theta}-y_t\right)^2}$. SGD with CS-PR to minimize $f(\theta)$ can be given as below:
\begin{subequations}\label{eq:linreg}
\theta_t&=\theta_{t-1}-\alpha x_t\left(\ip{x_t,\theta_{t-1}} -y_t\right)\\
\thh_t&=\frac{1}{t+1}\sum_{s=0}^t \theta_t
\end{subequations}
Here $\nabla f(\theta_{t-1})=x_t\left(\ip{x_t,\theta_{t-1} -y_t}\right)$ is the \emph{stochastic gradient} of the objective $f(\theta)$. It follows from the $i.i.d$ assumption on $(x_t,y_t)$ that $A_t=x_t x_t^\top$ and $b_t=y_t x_t$ are also $i.i.d$, and with this identification one can see that \eqref{linreg} is of the form in \eqref{conststep}. In this case $A_P$ is SPD.
In the linear regression setting, the error of a parameter $\theta$ is measured as the difference between the loss at the said parameter and the loss at the optimum, which has the following form
\begin{align}\label{eq:mselinreg}
f(\theta)-f(\ts)=(\theta-\ts)^\top A_P (\theta-\ts)=\EE{\norm{\theta-\ts}^2_{A_P}}
\end{example}
In this setting, \cite{bach} show that the variance term in the MSE is at most $O(\frac{C}{t})$, where $C$ does not depend on the condition number. This ``remarkable'' result is due to the following facts: $(i)$ $A_P$ in this case is symmetric positive definite, $(ii)$ the MSE in \eqref{eq:mselinreg} is measured with respect to $A_P$, $(iii)$ the noise $\epsilon_t=y_t-\ip{\ts,x_{t-1}}$ is structured i.e., $\EE{\epsilon_t^2 x_t x_t^\top}\leq R A_P$ for some constant $R>0$.
At the same time, the bias term decays at a rate $\frac{C'}{t^2}$, where the constant $C'$ contains the factor $(\theta_0-\ts)^\top A^{-1}_P (\theta_0-\ts)$ and is dependent on the condition number.

### 2017-05-08 intro.tex
updated version

LSA algorithms are useful in machine learning and reinforcement learning (RL) applications, where the aim is to compute a $\ts\in \R^d$ using the noisy data sequence $(b_t,A_t)$. Specific examples include the stochastic gradient descent algorithm for the problem of linear predition with quadratic loss, and the temporal difference class (TD) of learning algorithms in RL.
An additional feature in these class of pplications is that $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be obtained in $O(d)$, which is attractive due to the cheap per time step computational requirement.\par

@@

 We only assume that $\EE{A_t}$ is Hurwitz, i.e., all its eigenvalues have positive real parts. This is the condition used in prior works by \cite{polyak-judisky,borkarbook}. Eventhough our proof technique resembles that by \cite{bach}, in addition, we make use of Jordon/Cyclic decomposition in a critical way to handle the general Hurwitz assumption. Our results are applicable to wide variety of TD algorithms such as TD(0), GTD/GDT2, TDC.
Further, the error dynamics has a simple interpretation in terms of behaviour of the sum of terms in a geometric series, and hence offers a simple and unified understanding of the various TD algroithms.

>>>>

LSA algorithms are useful in machine learning and reinforcement learning (RL) applications, where the aim is to compute a $\ts\in \R^d$ using the noisy data sequence $(b_t,A_t)$ using cheap per iteration computations. Specific examples include the stochastic gradient descent algorithm for the problem of linear predition with quadratic loss, and the temporal difference class (TD) of learning algorithms in RL.\par
%An additional feature in these class of pplications is that $A_t$ turns out to be a rank-$1$ matrix and $A_t\theta_{t-1}$ can be obtained in $O(d)$, which is attractive due to the cheap per time step computational requirement.\par

@@

 We only assume that $\EE{A_t}$ is Hurwitz, i.e., all its eigenvalues have positive real parts. This is the condition used in prior works by \cite{polyak-judisky,borkarbook}. Eventhough our proof technique resembles that by \cite{bach}, in addition, we make use of Jordon/Cyclic decomposition in a critical way to handle the general Hurwitz assumption. Our results are applicable to wide variety of TD algorithms such as TD(0), GTD/GDT2, TDC. Further, the error dynamics has a simple interpretation in terms of behaviour of the sum of terms in a geometric series, and hence offers a simple and unified understanding of the various TD algroithms.
\paragraph{Organization:}

### 2017-05-08 prob.tex
updated version

\section{Problem Setup}
We consider the linear stochastic approximation (LSA) algorithm with constant step size and PR-averaging of the iterates.
\begin{subequations}\label{eq:lsa}
\begin{align}
\label{conststep}&\text{LSA:} &\theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),\\
\label{iteravg}&\text{PR-Average:} &\thh_t&=\frac{1}{t+1}{\sum}_{s=0}^{t}\theta_i,
\end{align}
\end{subequations}
where $\thh_t, \theta_t \in \C^{d}$. The algorithm updates a pair of parameters $\theta_t,\tb_t\in \C^{\dcd}$ incrementally, in discrete time steps $t=1,2,\dots$
based on data $b_t\in \C^d$, $A_t\in \C^{n\times n}$\footnote{In almost all practical instances, the iterates $\theta_t,\thh_t\in \R^d$ and $A_t\in \R^{\dcd}, b_t\in \R^d$. However, it serves well for our analysis to make the general assumption that these qunatites are complex valued.}. Here $\alpha>0$ is a positive step-size parameter; the only tuning parameter of the algorithm besides the
initial value $\theta_0$. The iterate $\theta_t$ is treated as an internal state of the algorithm, while $\th_t$ is the output at time step $t$. The update of $\theta_t$ alone is considered a form of constant stepsize LSA. Note that the computation of $\theta_t$ can also be performed in an incremental fashion using $O(n)$ storage.

Owning to the special form of $A_t$, sometimes the matrix-vector product $A_t \theta_{t-1}$
can be efficiently computed in $O(n)$ time.
This happens for example when $A_t$ is rank one; examples of this kind arising in reinforcemnet learning \cite{} and linear prediction \cite{} will be presented in a later sections (\Cref{sec:rl,sec:lp}). The significance of efficient computation of the matrix-vector product is that then an update of the algorithm
can be implemented in $O(n)$ time and $O(n)$ storage, which makes the algorithm particularly attractive in large-scale computations when $n$ is in the range of thousands or more.
We are interested in the behaviour of \eqref{eq:lsa} under the assumptions presented in \Cref{assmp:lsa} In what follows, we make use of the $\sigma$-field given by $\F_{t-1}=\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$.

@@

\textbf{Performance Metric:}
The ``aim'' of the updates in \eqref{eq:lsa} is to estimate $\ts$ based on the noisy versions $(A_t,b_t)$ of $(A_P,b)$.
There are various ways of quantifying the error $\eh_t \doteq \thh_t- \ts$ of estimating $\ts$ using $\thh_t$, but the magnitude
of the errors is usually measured in some weighted quadratic norm. The expected loss for time step $t$ then becomes
\begin{align}\label{eq:exploss}
L_t \doteq \EE{ \norm{\eh_t}_{C}^2 }\,,
where recall that for a SPD matrix $C$, $\norm{x}_C^2 = x^* C x$. Here, the choice of the matrix $C$ depends on the application.\\

>>>>

\section{Problem Setup}
We consider the linear stochastic approximation (LSA) algorithm with constant step-size (CS) and Polyak-Ruppert (PR) averaging of the iterates.
\begin{subequations}\label{eq:lsa}
\begin{align}
\label{conststep}&\text{LSA:} &\theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),\\
\label{iteravg}&\text{PR-Average:} &\thh_t&=\frac{1}{t+1}{\sum}_{s=0}^{t}\theta_i,
\end{subequations}
where\footnote{In almost all practical instances, the iterates $\theta_t,\thh_t\in \R^d$ and $A_t\in \R^{\dcd}, b_t\in \R^d$. However, it serves well for our analysis to consider these qunatites to be complex valued.} $\thh_t, \theta_t \in \C^{d}$. The algorithm updates a pair of parameters $\theta_t,\tb_t\in \C^{\dcd}$ incrementally, in discrete time steps $t=1,2,\dots$
based on data $b_t\in \C^d$, $A_t\in \C^{\dcd}$. Here $\alpha>0$ is a positive step-size parameter; the only tuning parameter of the algorithm besides the
initial value $\theta_0$. The iterate $\theta_t$ is treated as an internal state of the algorithm, while $\thh_t$ is the output at time step $t$. The update of $\theta_t$ alone is considered a form of constant step-size LSA.\par
Note that the computation of $\theta_t$ can also be performed in an incremental fashion using $O(d)$ storage. Owning to the special form of $A_t$, sometimes the matrix-vector product $A_t \theta_{t-1}$ can be efficiently computed in $O(d)$ time. This happens for example when $A_t$ is rank one; examples of this kind arising in reinforcemnet learning \cite{} and linear prediction \cite{} will be presented in a later sections (\Cref{sec:rl,sec:lp}). The significance of efficient computation of the matrix-vector product is that then an update of the algorithm
can be implemented in $O(d)$ time and $O(d)$ storage, which makes the algorithm particularly attractive in large-scale computations when $n$ is in the range of thousands or more.\par
We are interested in the behaviour of \eqref{eq:lsa} under the assumptions presented in \Cref{assmp:lsa}. In what follows, we make use of the $\sigma$-field given by $\F_{t-1}=\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$.

@@

%For the rest of the paper we denote the errors of the internal and the output variables of \eqref{eq:lsa} at time $t$ by $e_t\eqdef \theta_t-\ts$ and $\eh_t\eqdef \thh_t-\ts$ respectively.
\paragraph{Performance Metric:}  We are interested in the behavior of the mean squared error (MSE) at time $t$ given\footnote{ We note here that $\EE{\norm{\thh_t-\ts}_C^2}$ which measures the MSE in the quadratic norm with respect to a positive definite matrix $C\succ 0$. However, unless $C$ has some special structure, it is not possible to exploit this generality in our bounds. Further, when data is bounded $\norm{C}^2$ is also bounded and it is straightforward to see that $\EE{\norm{\thh_t-\ts}}^2_C\leq \norm{C}^2\EE{\norm{\thh_t-\ts}^2}$. So, for the purpose of stating our results we will not use the general quadratic norm, however, will use it when we discuss the work by \cite{bach} in the linear regression setting, where its usage is relevant.} by $\EE{\norm{\thh_t-\ts}^2}$. We wish to understand how the MSE is affected by the choice of $\alpha$ and the spectral properties of $A_P$. Another aspect of interest is to understand the contribution of the \emph{bias} (caused by the initial condition $\theta_0$) and the \emph{variance} (caused by $\sigma^2_P,\sigma^2_b$) to the MSE.\par
\begin{comment}
